---
title: "NJ OAG Use of force data cleaning workbook"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NJ OAG Use of force data cleaning workbook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(magrittr)
library(purrr)
library(stringr)
library(tidyr)
library(njoaguof)
```

## Set up

Load the Use of Force raw data.

```{r}
data("use_of_force_raw")
uof_raw <- use_of_force_raw
```

We will also need county and municipality data from the US Census, gathered using the `tidycensus` package:
```{r}
data("census_counties")
data("census_municipalities")
```

Some regular expressions we will use repeatedly

```{r}
trailing_comma_regex <- "(?<=.),?$"
sep_comma_no_space <- ",(?! )"
sep_comma_space_no_paren <- r"(,(?![^(]*\)) )"
```

## Pre-validation

Before we start parsing, we will verify some things.

### Column names

First, here are the columns we are expecting.

```{r}
raw_names <- c(
  "form_id"                          ,"County2",
  "agency_name3"                     ,"officer_name",
  "Officer_Name2"                    ,"Officer_Name_Agency",
  "INCIDENTID"                       ,"report_number",
  "incident_case_number"             ,"IncidentDate1",
  "incident_date"                    ,"IncidentDate1_old",
  "other_officer_involved"           ,"officer_in_uniform",
  "incident_municipality"            ,"indoor_or_outdoor",
  "incident_weather"                 ,"video_footage",
  "video_type"                       ,"incident_lighting",
  "location_type"                    ,"incident_type",
  "contact_origin"                   ,"planned_contact",
  "officer_age"                      ,"officer_race",
  "officer_rank"                     ,"officer_gender_fill",
  "OffInjuryType"                    ,"officer_injuries_injured",
  "OFFMEDTREAT2"                     ,"officer_hospital_treatment",
  "TotalSubInjuredIncdient"          ,"SubjectInjuredInIncident",
  "SubjectInjuredPrior"              ,"PerceivedCondition",
  "SubActions"                       ,"SubResist",
  "SubMedicalTreat"                  ,"SubjectInjuries",
  "SubjectsArrested"                 ,"ReasonNotArrest",
  "subject_type"                     ,"SubectsAge",
  "SubjectRace"                      ,"SubjectsGender",
  "TypeofForce"                      ,"Incident_date1",
  "incident_lighting2"               ,"IncidentYear"
)
stopifnot(all.equal(names(uof_raw), raw_names))
```

### Levels

We will use this function to check the levels of multi-valued fields.

```{r}
check_list_levels <- function(tbl,
                              column,
                              separating_regex,
                              levels,
                              other_acceptable_values = character(0)) {
  message("Checking for unmatched values in ",
          rlang::as_label(rlang::enquo(column)))
  unmatched_values <- tbl %>%
    dplyr::pull({{column}}) %>%
    stringr::str_split(separating_regex) %>%
    unlist() %>% unique() %>%
    setdiff(levels) %>% setdiff(other_acceptable_values)
  if (length(unmatched_values) != 0) {
    message(paste(unmatched_values, collapse = " "))
    stop()
  }
}
```

And here are all the levels.

```{r}
subject_type_levels <-
  c("Person", "Animal", "Other", "Unknown Subject(s)")

gender_levels <-
  c("Male",
    "Female",
    "Gender Non-Conforming/X",
    "Other")

incident_type_levels <- c(
  "Aggressive/Injured Animal",
  "Assault",
  "Assisting another officer",
  "Burglary",
  "Distribution of CDS",
  "Disturbance (drinking, fighting, disorderly)",
  "Domestic",
  "Eluding",
  "Medical Emergency",
  "MV Accident/Aid",
  "MV/Traffic Stop",
  "Pedestrian Stop",
  "Possession of CDS",
  "Potential Mental Health Incident",
  "Report of Gunfire",
  "Robbery",
  "Subject with a gun",
  "Subject with other weapon",
  "Suspicious person",
  "Terroristic Threats",
  "Theft/Shoplifting",
  "Trespassing",
  "Wanted Person",
  "Welfare Check",
  "Other"
)

race_levels <- c(
  "American Indian",
  "Asian",
  "Asian/Pacific Islander",
  "Black",
  "Hispanic",
  "Native Hawaiian or other Pacific Islander",
  "Pacific Islander",
  "Two or more races",
  "White",
  "Other"
)

weather_levels <- c("Clear", "Cloudy", "Fog", "Rain", "Snow/Sleet/Ice")

video_type_levels <- c(
  "Body Worn",
  "CED Camera",
  "Cell Phone",
  "Commercial Building",
  "Motor Vehicle",
  "Residential/Home",
  "Station House",
  "Other",
  "Unknown"
)

lighting_levels <- c("Artificial", "Darkness", "Dawn/Dusk", "Daylight")

location_type_levels <- c(
  "Alcohol Establishment(bar, club, casino)",
  "Business",
  "Court House",
  "Hospital",
  "Jail/Prison",
  "Police Station",
  "Residence",
  "Restaurant",
  "School",
  "Street",
  "Other"
)

contact_origin_levels <- c(
  "Citizen Initiated",
  "Dispatched",
  "Officer Initiated",
  "Pre-Planned Contact"
)

planned_contact_levels <- c(
  "Arrest",
  "Judicial Order Service (TRO, FRO, etc.)",
  "Prisoner Transfer",
  "Processing",
  "Search Warrant Execution",
  "Other"
)

officer_injury_type_levels <- c(
  "Abrasion/Laceration/Puncture",
  "Chest pains/shortness of breath",
  "Complaint of pain",
  "Concussion",
  "Contusion/bruise",
  "Fracture/dislocation",
  "Gunshot wound",
  "Other",
  "Unknown"
)

officer_medical_treatment_levels <- c(
  "EMS on scene",
  "Hospital",
  "Officer Administered First Aid",
  "Refused",
  "Urgent Care"
)

perceived_condition_levels <- c(
  "No unusual condition noted",
  "Other unusual condition noted",
  "Potential Mental Health Incident",
  "Under influence of alcohol/drugs/both"
)

subject_action_levels <- c(
  "Attack with Blunt object",
  "Attack with Bodily fluids",
  "Attack with Edge Weapon",
  "Attack with Hands, fists, legs",
  "Attack with Motor Vehicle",
  "Attack with Other Weapon",
  "Attempt to commit crime",
  "Attempt to destroy evidence",
  "Attempt to escape from Custody",
  "Attempt to self-harm",
  "Biting",
  "Failure to Disperse",
  "Fired Gun",
  "Kick",
  "Other Attack",
  "Other Threat",
  "Prevent harm to another",
  "Push or shove",
  "Resisted arrest/police officer control",
  "Spitting",
  "Strike with open hand, fist, or elbow",
  "Threat to Kick",
  "Threat to Push or shove",
  "Threat to Strike with open hand, fist, or elbow",
  "Threat with Blunt object",
  "Threat with Bodily Fluids",
  "Threat with Edge Weapon",
  "Threat with Gun",
  "Threat with Hands, fists, legs",
  "Threat with Motor vehicle",
  "Threat with Other Weapon",
  "Verbal/Fighting stance Threat"
)

subject_resistance_levels <- c(
  "Active Assailant",
  "Active Resistor",
  "Aggressive resistance(attempt to attack or harm)",
  "Attempt to flee",
  "Dead-weight tactics(going limp)",
  "Non-response (consciously ignoring)",
  "Passive Resistor",
  "Resistive tension(stiffening, tightening muscles)",
  "Threatening Assailant",
  "Verbal",
  "Other"
)

subject_medical_treatment_levels <- c(
  "EMS on scene",
  "Hospital",
  "Mental Health Facility",
  "Officer Administered First Aid",
  "Refused",
  "Urgent Care",
  "Unknown",
  "Not Provided"
)

subject_injury_levels <- c(
  "Abrasion/Laceration/Puncture",
  "Chest pains/shortness of breath",
  "Complaint of pain",
  "Concussion",
  "Contusion/bruise",
  "Fracture/dislocation",
  "Gunshot wound",
  "No Injury",
  "Other",
  "Unknown",
  "Not Provided"
)

force_type_levels <- c(
  "Canine bit (apprehension)",
  "Canine bit (spontaneous)",
  "CED Spark Display",
  "Chokehold, Carotid artery restraint",
  "Compliance hold with impact weapon- not a strike",
  "CS Gas",
  "Discharged Chemical at",
  "Discharged Firearm at",
  "High Volume OC Spray",
  "Kneeling on Chest, Back",
  "Pointing Firearm",
  "Struck",
  "Used arm bar on",
  "Used arms",
  "Used arms/hands",
  "Used CED on",
  "Used fists/punch",
  "Used head",
  "Used legs/kicks",
  "Used Less-lethal device on",
  "Used pressure points on",
  "Used take down on",
  "Other"
)

reason_not_arrested_levels <- c(
  "Already in Custody",
  "Deceased",
  "Insufficient Probable Cause- includes continuing investigation",
  "Medical/Mental Health Incident",
  "No Probable Cause- Crime Unfounded",
  "No Probable Cause- Subject Not Involved",  
  "Subject Fled",
  "Other"
)

county_levels <- census_counties %>% pull(county)
```


Now we will check that the multi-value fields contain only these levels.

```{r}
uof_raw_trimmed <- uof_raw %>%
  dplyr::mutate(across(
  where(is.character),
  ~ stringr::str_replace(., trailing_comma_regex, "")
))

check_list_levels(
  uof_raw_trimmed,
  subject_type,
  sep_comma_no_space,
  subject_type_levels,
  ""
)
check_list_levels(
  uof_raw_trimmed,
  SubjectsGender,
  sep_comma_no_space,
  gender_levels,
  c("", "Not Provided")
)
check_list_levels(
  uof_raw_trimmed,
  SubjectRace,
  sep_comma_no_space,
  race_levels,
  c("Am. Indian", "Black or African American", "", "Not Provided")
)
check_list_levels(
  uof_raw_trimmed,
  incident_weather,
  sep_comma_space_no_paren,
  weather_levels,
  c("N/A", "")
)
check_list_levels(
  uof_raw_trimmed,
  video_type,
  sep_comma_space_no_paren,
  video_type_levels,
  c("[]", "")
)
check_list_levels(
  uof_raw_trimmed,
  incident_lighting,
  sep_comma_space_no_paren,
  lighting_levels
)
check_list_levels(
  uof_raw_trimmed,
  location_type,
  sep_comma_space_no_paren,
  location_type_levels,
  ""
)
check_list_levels(
  uof_raw_trimmed,
  incident_type,
  sep_comma_space_no_paren,
  incident_type_levels,
  ""
)
check_list_levels(
  uof_raw_trimmed,
  contact_origin,
  sep_comma_space_no_paren,
  contact_origin_levels,
  ""
)
check_list_levels(
  uof_raw_trimmed,
  planned_contact,
  sep_comma_space_no_paren,
  planned_contact_levels,
  c("", "Other ")
)
check_list_levels(
  uof_raw_trimmed,
  OffInjuryType,
  sep_comma_space_no_paren,
  officer_injury_type_levels,
  c("No Injury", "Not Provided", "")
)
check_list_levels(
  uof_raw_trimmed,
  OFFMEDTREAT2,
  sep_comma_space_no_paren,
  officer_medical_treatment_levels,
  c("Not Provided", "")
)
check_list_levels(
  uof_raw_trimmed,
  PerceivedCondition,
  sep_comma_no_space,
  perceived_condition_levels,
  c("Not Provided", "")
)
check_list_levels(
  uof_raw_trimmed,
  SubActions,
  sep_comma_no_space,
  subject_action_levels,
  ""
)
check_list_levels(
  uof_raw_trimmed,
  SubResist,
  sep_comma_no_space,
  subject_resistance_levels,
  ""
)
check_list_levels(
  uof_raw_trimmed,
  SubMedicalTreat,
  sep_comma_no_space,
  subject_medical_treatment_levels,
  ""
)
check_list_levels(
  uof_raw_trimmed,
  SubjectInjuries,
  sep_comma_no_space,
  subject_injury_levels,
  ""
)
# For TypeOfForce, we have to use an ad-hoc separator expression,
# because of internal commas in two levels.
sep_comma_space_force_special <- r"(, (?!Back|Carotid))"
check_list_levels(
  uof_raw_trimmed,
  TypeofForce,
  sep_comma_space_force_special,
  force_type_levels,
  ""
)
# There is one value here that we want to normalize --
# The value "NoProbableCause-SubjectNotInvolved. We will do that
# later, when we build the incident_subject_reason_not_arrested
# table.
 check_list_levels(
   uof_raw_trimmed,
   ReasonNotArrest,
   sep_comma_no_space,
   reason_not_arrested_levels,
  c("Not Provided", "", "NoProbableCause-SubjectNotInvolved")
)
```

Finally, we check the county field.  There are two non-county values which we 
are expecting (and will map to `NA`): `"Other"` and `"NJSP"`.

```{r}
stopifnot(all(
  setdiff(uof_raw %>% dplyr::pull(County2) %>% paste0(" County"),
        county_levels) %in%
    c("Other County", "NJSP County")
  )
)
```

## `subject` table

```{r}
subject <- uof_raw %>%
  select(form_id,
         SubjectsArrested,
         subject_type,
         SubectsAge,
         SubjectRace,
         SubjectsGender) %>%
  mutate(across(
    where(is.character),
    ~ str_replace(., trailing_comma_regex, "")
  )) %>%
  filter(if_any(-form_id, ~ . != ""))
  
max_subjects <- 
  subject$SubjectsArrested %>% 
  map_int(str_count, ",") %>% 
  max() + 1


subject <- subject %>% 
  separate(SubjectsArrested, 
           paste0("arrested__", 1:max_subjects), 
           ",",
           fill="right") %>%
  separate(subject_type, 
           paste0("type__", 1:max_subjects), 
           ",",
           fill="right") %>%
  separate(SubectsAge, 
           paste0("age__", 1:max_subjects), 
           ",",
           fill="right") %>%
  separate(SubjectRace, 
           paste0("race__", 1:max_subjects), 
           ",",
           fill="right") %>%
  separate(SubjectsGender, 
           paste0("gender__", 1:max_subjects), 
           ",",
           fill="right") %>%
  pivot_longer(cols = -form_id,
               names_to="column",
               values_to="value",
               values_drop_na=TRUE) %>%
  separate(column, c("column", "index"), "__") %>%
  pivot_wider(names_from="column", values_from="value") 

```

Now we will clean up this table, column by column.

#### `index` and `arrested`

`index` is the index into the subject list. It should be an integer. `arrested` is a boolean.

```{r}
subject <- subject %>% 
  mutate(index=as.integer(index),
         arrested=as.logical(arrested)
  )

```

#### `type` and `gender`

We will convert these to factors.

```{r}
subject <- subject %>% 
  mutate(type=factor(type, levels=subject_type_levels),
         gender=factor(gender, levels=gender_levels))
```

#### `age`

This field is not currently strictly numeric -- it also contains values `"Unknown"` and `"Juvenile"`. We will add a boolean column, `juvenile` and convert `age` to an integer. The value of `"Unknown"` will map to `NA` in both columns.

```{r}
as_integer_or_na <- function(x) suppressWarnings(as.integer(x))
subject <- subject %>% 
  mutate(juvenile=case_when(
    age=="Juvenile" ~ TRUE,
    !is.na(as_integer_or_na(age)) ~ FALSE
  )) %>%
  mutate(age=as_integer_or_na(age))
```

#### `race`

We need to normalize some names before converting to a factor.

```{r}
subject <- subject %>%
  mutate(
    race = case_when(
      race == "Black or African American" ~ "Black",
      race == "Am. Indian"                ~ "American Indian",
      TRUE                                ~ race
    ),
    race=factor(race, levels=race_levels)
  )
```

#### Cleanup and finish

Finally, we reorder the columns.

```{r}
subject <- subject %>% 
  select(form_id, index, arrested, type, age, juvenile, race, gender)
```

```{r purl=FALSE}
subject
```

## Incident set membership tables

These tables are built from columns that contain value-lists with the following properties:

1.  For each column, the set of values is drawn from a finite list
2.  For each field, there are no repeated values.
3.  The order of the list elements does not matter.

For each of these columns, we will create a new table with one row per value, using this function:

```{r}
### table should have two columns: form_id and list_col
make_set_membership_table <- function(table, levels, separating_regex = ",") {
  table <- table %>%
    mutate(list_col = str_replace(list_col, trailing_comma_regex, "")) %>%
    filter(list_col != "")

  max_values <- table$list_col %>%
    map_int(str_count, separating_regex) %>%
    max() + 1

  table %>%
    separate(list_col,
             paste0("list_col__", 1:max_values),
             separating_regex,
             fill="right"
    ) %>%
    pivot_longer(cols=-form_id,
                 names_to="column",
                 values_to="value",
                 values_drop_na = TRUE) %>%
    mutate(value=factor(str_trim(value), levels=levels)) %>%
    filter(!is.na(value)) %>%
    select(form_id, value)
}
```

### `incident_weather`

```{r}
incident_weather <- uof_raw %>% 
  select(form_id, list_col=incident_weather) %>%
  make_set_membership_table(weather_levels, 
                            sep_comma_space_no_paren) %>%
  rename(weather=value)
```

### `incident_video_type`

```{r}
incident_video_type <- uof_raw %>% 
  select(form_id, list_col=video_type) %>%
  make_set_membership_table(video_type_levels,
                            sep_comma_space_no_paren) %>%
  rename(video_type=value)
```

### `incident_lighting`

```{r}
incident_lighting <- uof_raw %>% 
  select(form_id, list_col=incident_lighting) %>%
  make_set_membership_table(lighting_levels,
                            sep_comma_space_no_paren) %>%
  rename(lighting=value)
```

### `incident_location_type`

```{r}
incident_location_type <- uof_raw %>% 
  select(form_id, list_col=location_type) %>%
  make_set_membership_table(location_type_levels, 
                            sep_comma_space_no_paren) %>%
  rename(location_type=value)
```

### `incident_type`

```{r}
incident_type <- uof_raw %>%
  select(form_id, list_col=incident_type) %>%
  make_set_membership_table(incident_type_levels,
                            sep_comma_space_no_paren) %>%
  rename(type=value)

```

### `incident_contact_origin`

```{r}
incident_contact_origin <- uof_raw %>%
  select(form_id, list_col=contact_origin) %>%
  make_set_membership_table(contact_origin_levels,
                            sep_comma_space_no_paren) %>%
  rename(contact_origin=value)
```

### `incident_planned_contact`

```{r}
incident_planned_contact <- uof_raw %>%
  select(form_id, list_col=planned_contact) %>%
  make_set_membership_table(planned_contact_levels,
                            sep_comma_space_no_paren) %>%
  rename(planned_contact=value)
```

### `incident_officer_injury_type`

This is pulled from the field `OffInjuryType`, which is of mixed type. The field may be equal to `"No Injury"` or to `"Not Provided"`. Otherwise, it is value-list column, using the levels below. When it is a value-list column, we will use it to populate the `incident_officer_injury_type` table. The other values will be reflected in the `incident` table.

```{r}

incident_officer_injury_type <- uof_raw %>%
  select(form_id, list_col=OffInjuryType) %>%
  make_set_membership_table(officer_injury_type_levels,
                            sep_comma_space_no_paren) %>%
  rename(officer_injury_type=value)
```

### `incident_officer_medical_treatment`

```{r}
incident_officer_medical_treatment <- uof_raw %>%
  select(form_id, list_col=OFFMEDTREAT2) %>%
  make_set_membership_table(officer_medical_treatment_levels,
                            sep_comma_space_no_paren) %>%
  rename(officer_medical_treatment=value)
```

## Messy relationship tables

These tables are built from columns that contain value-lists that are "messy" in the following sense. Each column contains value-lists with repeated values -- because the values apply to individual subjects -- but there is no reliable way to assign individual list elements to individual subjects. For example, `form_id` `20221` has two subjects, but field `SubActions` has three values:

```{r purl=FALSE}
uof_raw %>% 
  filter(form_id==20221) %>%
  mutate(SubActions=str_replace(SubActions, ",?$", "")) %>%
  select(form_id, subject_type, SubActions) %>%
  separate(SubActions, into=paste0("SubActions_", 1:3), ",") %>% 
  glimpse()
```

The value `"Resisted arrest/police officer control"` appears twice because it applies to both subjects. But `"Prevent harm to another"` appears only once, and it not clear if it applies to the first subject or to the second one.

We will create a separate table for each of these columns, assigning all of the values -- including the repeated values -- to the incident. There will also be an index column, so that the list position can be reconstructed if necessary.

```{r}
make_messy_relationship_table <- function(table, levels, separating_regex) {
  table <- table %>%
    mutate(list_col = str_replace(list_col, trailing_comma_regex, "")) %>%
    filter(list_col != "")

  max_values <- table$list_col %>%
    map_int(str_count, separating_regex) %>%
    max() + 1

  table %>%
    separate(list_col,
             paste0("value__", 1:max_values),
             separating_regex,
             fill="right"
    ) %>%
    pivot_longer(cols = -form_id,
                 names_to="column",
                 values_to="value",
                 values_drop_na=TRUE) %>%
    separate(column, c("column", "index"), "__") %>%
    mutate(index=as.integer(index)) %>%
    pivot_wider(names_from="column", values_from="value") %>%
    mutate(value=factor(str_trim(value), levels=levels)) %>%
    filter(!is.na(value))
}
```

### `incident_subject_perceived_condition`

```{r}
incident_subject_perceived_condition <- uof_raw %>%
  select(form_id, list_col=PerceivedCondition) %>%
  make_messy_relationship_table(perceived_condition_levels,
                                sep_comma_no_space) %>%
  rename(perceived_condition=value)
```

### `incident_subject_action`

```{r}

incident_subject_action <- uof_raw %>% 
  select(form_id, list_col=SubActions) %>%
  make_messy_relationship_table(subject_action_levels,
                                sep_comma_no_space) %>%
  rename(subject_action=value)
```

### `incident_subject_resistance`

```{r}
incident_subject_resistance <- uof_raw %>% 
  select(form_id, list_col=SubResist) %>%
  make_messy_relationship_table(subject_resistance_levels,
                                sep_comma_no_space) %>%
  rename(subject_resistance=value)
```

### `incident_subject_medical_treatment`

```{r}
incident_subject_medical_treatment <- uof_raw %>% 
  select(form_id, list_col=SubMedicalTreat) %>%
  make_messy_relationship_table(subject_medical_treatment_levels,
                                sep_comma_no_space) %>%
  rename(subject_medical_treatment=value)
```

### `incident_subject_injury`

```{r}
incident_subject_injury <- uof_raw %>% 
  select(form_id, list_col=SubjectInjuries) %>%
  make_messy_relationship_table(subject_injury_levels,
                                sep_comma_no_space) %>%
  rename(subject_injury=value)
```

### `incident_subject_force_type`

```{r}
incident_subject_force_type <- uof_raw %>% 
  select(form_id, list_col=TypeofForce) %>%
  make_messy_relationship_table(force_type_levels,
                                sep_comma_space_no_paren) %>%
  rename(force_type=value)
```

### `incident_subject_reason_not_arrested`

One value in this field needs to be normalized: We will replace
`"NoProbableCause-SubjectNotInvolved"` with `"No Probable Cause- Subject Not Involved"`.

```{r}
incident_subject_reason_not_arrested <- uof_raw %>%
  mutate(ReasonNotArrest = str_replace_all(
    ReasonNotArrest,
    "NoProbableCause-SubjectNotInvolved",
    "No Probable Cause- Subject Not Involved"
  )) %>%
  select(form_id, list_col="ReasonNotArrest") %>%
  make_messy_relationship_table(reason_not_arrested_levels,
                                sep_comma_no_space) %>%
  rename(reason_not_arrested=value)
```

## `incident` table

### Set up

First, we select the columns we will need from the raw data.

```{r}
incident <- uof_raw %>%
  select(
    form_id,
    agency_county = County2,
    agency_name = agency_name3,
    officer_name = Officer_Name2,
    officer_name_id = officer_name,
    report_number,
    incident_case_number,
    incident_date_1 = IncidentDate1,
    incident_municipality,
    indoor_or_outdoor,
    video_footage,
    officer_age,
    officer_race,
    officer_rank,
    officer_gender = officer_gender_fill,
    officer_injured = officer_injuries_injured,
    subject_injured_count = TotalSubInjuredIncdient
  )
```

### Fields not selected

Before we clean up these fields, let us review the fields we are omitting.

#### `INCIDENTID` and `Officer_Name_Agency`

These are synthetic fields composed of other fields, so we do not need them.

```{r}
stopifnot(
  0 ==
    uof_raw %>%
    filter(
      INCIDENTID !=
        glue::glue("{County2}-{agency_name3}-{incident_case_number}")
    ) %>%
    filter(
      INCIDENTID !=
        glue::glue("{County2}-{agency_name3}- {incident_case_number}")
    ) %>%
    filter(
      INCIDENTID !=
        glue::glue("{County2}-{agency_name3}-\t{incident_case_number}")
    ) %>%
    filter(
      INCIDENTID !=
        glue::glue("{County2}-{agency_name3}-  {incident_case_number}")
    ) %>%
    nrow(),
  0 ==
    uof_raw %>%
      filter(
        Officer_Name_Agency !=
          str_to_upper(glue::glue("{agency_name3}-{Officer_Name2}"))
      ) %>%
      filter(
        Officer_Name_Agency !=
          str_to_upper(glue::glue("{agency_name3}-\t\t{Officer_Name2}"))
      ) %>%
      filter(
        Officer_Name_Agency !=
          str_to_upper(glue::glue("{agency_name3}-\t\t\t{Officer_Name2}"))
      ) %>%
      nrow()
)
```

#### `IncidentDate1_old` `Incident_date1` and `incident_date`

These fields are always identical to `IncidentDate1`. 
For some reason, `haven` didn't catch that `Incident_date1` is an SPSS date, so we will do that conversion manually here.

```{r}
stopifnot(
  0 ==
    uof_raw %>% 
    mutate(IncidentDate1_old = as.Date(IncidentDate1_old / 86400, origin =
                                      "1582-10-14")) %>%
    mutate(IncidentDate1_old = lubridate::floor_date(IncidentDate1_old,
                                                   unit="day")) %>%
    filter(IncidentDate1 != IncidentDate1_old) %>% nrow(),
  0 ==
    uof_raw %>%
    mutate(Incident_date1 = as.Date(Incident_date1)) %>%
    filter(IncidentDate1 != Incident_date1) %>% nrow(),
  0 ==
    uof_raw %>% filter(incident_date != IncidentDate1) %>% nrow()
)
```

#### `other_officer_involved` and `officer_in_uniform`

These fields are always `NA`:

```{r}
stopifnot(0 == uof_raw %>% filter(!is.na(other_officer_involved)) %>% nrow(),
          0 == uof_raw %>% filter(!is.na(officer_in_uniform)) %>% nrow())
```

#### `incident_lighting2`

This field is always `1`:

```{r}
stopifnot(0 == uof_raw %>% filter(incident_lighting2 != 1) %>% nrow())
```

#### `incident_weather`, `video_type`, `incident_lighting`, `location_type`, `incident_type`, `contact_origin`, `planned_contact`, `OFFMEDTREAT2`, `PerceivedCondition`,`SubActions`,`SubResist`, `SubMedicalTreat`, `SubjectInjuries`, `TypeofForce`, and `ReasonNotArrest`

These fields are handled in other tables.

#### `OffInjuryType`

We did not capture the `"No Injury"` or empty-string value in the `incident_officer_injury_type` table, but that value is redundant with `officer_injuries_injured` which is `"1"` precisely when `OffInjuryType`is `"NoInjury"` or the empty string.

```{r}
stopifnot(
  0 == uof_raw %>%
    select(OffInjuryType, officer_injuries_injured) %>%
    mutate(NoInjury = str_detect(OffInjuryType, "No Injury") |
             OffInjuryType == "") %>%
    filter(NoInjury == (officer_injuries_injured == "1")) %>%
    nrow()
)
```

#### `SubjectInjuredInIncident` and `SubjectInjuredPrior`

This field is identical with `TotalSubInjuredIncdient`

```{r}
stopifnot(
  0 ==
    uof_raw %>% filter( 
      (TotalSubInjuredIncdient != SubjectInjuredInIncident),
        !(TotalSubInjuredIncdient == 0 & SubjectInjuredInIncident == "")
    ) %>%
    nrow(),
  0 ==
    uof_raw %>% filter(
      (TotalSubInjuredIncdient != SubjectInjuredPrior),
        !(TotalSubInjuredIncdient == 0 & SubjectInjuredPrior == "")
    ) %>%
    nrow()
)
```

#### `SubjectsArrested`, `subject_type`, `SubectsAge`, `SubjectRace`, and `SubjectsGender`

These fields are reflected in the `subject` table.

#### `IncidentYear`incident_date

`IncidentYear` is redundant with `IncidentDate1`.

```{r}
stopifnot(
  all(
    uof_raw$IncidentYear == lubridate::year(uof_raw$IncidentDate1)
  )
)
```

### `incident` fields

#### `agency_county`

Switch to using the census county names. In particular, the non-counties `"NJSP"` and `"Other"` will be mapped to `NA`.

```{r}
incident <- incident %>% 
  mutate(agency_county=paste0(agency_county, " County")) %>%
  mutate(agency_county = factor(agency_county, county_levels))
```

#### `agency_name`
There are over 400 distinct values in this field, at least 19 of which appear exactly once, so we will not use a factor here.

In order to reconcile the agency names in this dataset with the agency
names in the NJ OAG Law Enforcement Officer Diversity dataset, we will make some modifications.
```{r}
incident <- incident %>%
  mutate(
    agency_name = case_when(
      agency_name == "Burlington County Sheriffs Office" ~ "Burlington Co Sheriffs Office",
      agency_name == "Mercer Co Prosecutor Off" ~ "Mercer Co Prosecutors Office",
      agency_name == "Hudson Co ProsecutorOff-S/Force" ~ "Hudson Co Prosecutors Office",
      agency_name == "Somerset Co Sheriffs Dept" ~ "Somerset Co Sheriffs Office",
      agency_name == "Middle Twsp PD" ~ "Middle Twp PD",
      agency_name == "State Police" ~ "New Jersey State Police",
      agency_name == "Rutgers Univ Police" ~ "Rutgers University PD",
      agency_name == "Division Of Fish And Wildlife" ~ "NJ Division Of Fish And Wildlife",
      agency_name == "NJ State Human Services Police" ~ "NJ Department Of Human Services",
      agency_name == "Division of Criminal Justice" ~ "NJ Division of Criminal Justice",
      agency_name == "Park Police" ~ "New Jersey State Park Police",
      agency_name == "Toms River Township" ~ "Toms River Twp PD",
      agency_name == "Hardyston Twp Police Dept" ~ "Hardyston Twp PD",
      agency_name == "Linden Police Department" ~ "Linden PD",
      TRUE ~ agency_name
    )
  )
```


#### `officer_name` and `officer_name_id`

The `officer_name` associated to a single `officer_name_id` is not always consistent in spelling and capitalization. For example:

```{r purl=FALSE}
incident %>% 
  group_by(officer_name_id) %>% 
  summarise(distinct_names=n_distinct(officer_name)) %>%
  filter(distinct_names>1) %>%
  left_join(select(incident, officer_name, officer_name_id), 
            by="officer_name_id") %>%
  unique()
```

Therefore, we will choose the most common variant for each name id and apply that to all rows. We will save variant names in a new table, `officer_name_variants`.

```{r}
incident <- incident %>% 
  mutate(officer_name = na_if(officer_name, ""))

standard_names <- incident %>% 
  count(officer_name, officer_name_id) %>%
  group_by(officer_name_id) %>%
  slice(which.max(n)) %>%
  ungroup() %>%
  select(officer_name_id, officer_name)

officer_name_variants <- incident %>%
  select(officer_name_id, officer_name) %>%
  unique()

incident <- incident %>% 
  select(-officer_name) %>%
  left_join(standard_names, by="officer_name_id") 

rm(standard_names)
```

#### `report_number`, `incident_case_number`, and `incident_date_1`

We leave these fields as they are.

#### `incident_municipality`
This field is of 
the form `<municipality>, <county-name>`. We will break this field up into `incident_municipality` and `incident_municipality_county`.

```{r}
incident <- incident %>%
  separate(
    incident_municipality,
    c("incident_municipality", "incident_municipality_county"),
    sep = ",",
    fill = "right"
  ) %>%
  mutate(incident_municipality_county =
           str_trim(incident_municipality_county))
```

Next, we will build a lookup table, matching the unique municipality/county
values in the UOF dataset to the census municipality names.  This is done in
three steps.  
```{r}
unique_municipalities <- incident %>%
  select(incident_municipality_county, incident_municipality) %>%
  unique()
```

First step:  Exact matches:
```{r}
lookup_1 <- unique_municipalities %>%
  inner_join(census_municipalities,
             by=c("incident_municipality"="municipality_and_type",
                  "incident_municipality_county"="county")) %>%
  mutate(census_municipality=incident_municipality) %>%
  select(incident_municipality, incident_municipality_county, census_municipality)
```
Second step: Matches with the base name (excluding the municipality "type", such as "borough" or "township"):
```{r}
lookup_2 <- unique_municipalities %>%
  anti_join(lookup_1,
            by=c("incident_municipality", "incident_municipality_county")) %>%
  inner_join(census_municipalities,
             by=c("incident_municipality"="municipality",
                  "incident_municipality_county"="county")) %>%
  mutate(census_municipality=municipality_and_type) %>%
  select(incident_municipality, incident_municipality_county, census_municipality)
```
Third step:  The remaining municipalities require ad-hoc logic.
```{r}
lookup_3 <- unique_municipalities %>%
  anti_join(
    lookup_1,
    by = c("incident_municipality", "incident_municipality_county")
  ) %>%
  anti_join(
    lookup_2,
    by = c("incident_municipality", "incident_municipality_county")
  ) %>%
  mutate(
    census_municipality = case_when(
      # Name changes:  Dover, Ocean County -->           Toms River
      #                Washington, Mercer County -->     Robbinsville
      #                West Paterson, Essex County -->   Woodland Park
      #                South Belmar, Monmouth County --> Spring Lake
      incident_municipality_county == "Ocean County" &
        incident_municipality == "Dover" ~ "Toms River township",
      incident_municipality_county == "Mercer County" &
        incident_municipality == "Washington" ~ "Robbinsville township",
      incident_municipality_county == "Passaic County" &
        incident_municipality == "West Paterson" ~ "Woodland Park borough",
      incident_municipality_county == "Monmouth County" &
        incident_municipality == "South Belmar" ~ "Spring Lake borough",
      # "Princeton township" and "Princeton borough" merged to form just-plain
      # "Princeton" -- the only municipality in NJ without a type!
      incident_municipality_county == "Mercer County" &
        incident_municipality == "Princeton township" ~ "Princeton",
      # Cleanup miscellaneous idiosyncrasies.
      incident_municipality_county == "Essex County" &
        incident_municipality == "Village of South Orange" ~ "South Orange Village township",
      incident_municipality_county == "Essex County" &
        incident_municipality == "Caldwell Borough" ~ "Caldwell borough",
      incident_municipality_county == "Monmouth County" &
        incident_municipality == "Spring Lake Boro" ~ "Spring Lake borough",
      incident_municipality_county == "Somerset County" &
        incident_municipality == "Peapack & Gladstone" ~ "Peapack and Gladstone borough"
    )
  )
```

Now build the full lookup table:
```{r}
lookup <- rbind(lookup_1, lookup_2, lookup_3)
```
Check that we have no duplicates, and that the only municipality we are mapping
to `NA` is `"Other"`:
```{r}
stopifnot(0 ==
  lookup %>% count(incident_municipality, incident_municipality_county) %>%
    filter(n>1) %>% nrow()
)
stopifnot("Other" == 
            lookup %>% filter(is.na(census_municipality)) %>% pull(incident_municipality)
)
```

Now, finally, we can use the lookup table to tidy the `incident_municipality field`:
```{r}
incident <- incident %>% 
  left_join(lookup, 
            by = c("incident_municipality", "incident_municipality_county")) %>% 
  mutate(incident_municipality=census_municipality) %>%
  select(-census_municipality) 
```

Let's also use the county levels for `incident_municipality_county`:
```{r}
incident <- incident %>%
  mutate(incident_municipality_county =
         factor(incident_municipality_county, county_levels))
```

#### `indoor_or_outdoor`

This field could indicate one or the other, or both, though the order is not consistent:

```{r purl=FALSE}
incident %>% count(indoor_or_outdoor)
```

We will break this into a pair of booleans.

```{r}
incident <- incident %>%
  mutate(
    indoors = str_detect(indoor_or_outdoor, "Indoors"),
    outdoors = str_detect(indoor_or_outdoor, "Outdoors")
  ) %>%
  select(-indoor_or_outdoor)
```

#### `video_footage`.

We will correct the spelling of `"Unknown"` before using a factor.

```{r}
video_footage_levels <- c("Yes", "No", "Unknown")
incident <- incident %>%
  mutate(video_footage=str_replace(video_footage, "Unknow", "Unknown")) %>%
  mutate(video_footage=factor(video_footage, levels=video_footage_levels))
```

#### `officer_age`

We remove implausible ages and convert to an integer:

```{r}
incident <- incident %>%
  dplyr::mutate(officer_age = ifelse(dplyr::between(officer_age, 18, 67),
                                     as.integer(officer_age),
                                     NA)) 
```

#### `officer_race` and `officer_gender`

We perform the same transformations as we did on the `subject` `race` and `gender` columns.

```{r}
incident <- incident %>%
  mutate(
    officer_race = case_when(
      officer_race == "Black or African American" ~ "Black",
      officer_race == "Am. Indian"                ~ "American Indian",
      TRUE                                        ~ officer_race
    ),
    officer_race = factor(officer_race, levels = race_levels)
  )

incident <- incident %>% 
  mutate(officer_gender=factor(officer_gender, levels=gender_levels))

```

#### `officer_rank`

This field appears to be free text, and there are a lot of variations. We leave the column unchanged.

#### `officer_injured`

We will convert this to a boolean.

```{r}
incident <- incident %>% 
  mutate(officer_injured = (officer_injured=="1"))
```

#### `subject_injured_count`

We leave this field unchanged.

### Clean up and finish

#### Add a `subject_count` column

```{r}
incident <- incident %>%
  left_join(count(subject, form_id, name = "subject_count"),
            by = "form_id")
```

#### Order the columns

```{r}
incident <- incident %>%
  relocate(
    form_id,
    report_number,
    incident_case_number,
    incident_date_1,
    agency_county,
    agency_name,
    incident_municipality,
    incident_municipality_county,
    officer_name_id,
    officer_name,
    officer_age,
    officer_race,
    officer_rank,
    officer_gender,
    officer_injured,
    video_footage,
    indoors,
    outdoors,
    subject_count,
    subject_injured_count
  )
```

```{r purl=FALSE}
incident
```
